### 2011-06-02 docs/pubs/EIM-2011/application.tex
figures for: integration, metacat query, morpho annotation

\subsection{Data query}
For even more precise recall, the OBOE model can be partially \emph{materialized} during the query stage after which a data range filter can be applied. Different techniques are available for merging the Annotation with the data that it describes, but for performance reasons a hybrid approach has been adopted in which premliminary search results from a concept-based query are collated and only that subset is materialized. The Data Manager Library is conveniently used to load raw data objects because our corpus is described using EML and the Annotation syntax - both of which inform the correct use of the Data Manager query features. For any data attributes that match the concept query criteria, we then verify that those attributes contain data values within the range specifed by the initial semantic data query. Figure XX illustrates a semantic data query in which datasets recording organismal mass is greater than 0.5 grams are returned.

\subsection{Data integration}
The materialization routine used for semantic data queries laid the groundwork for our ongoing data integration pursuits.
In addition to inspecting the data for values within a range and returning the packages that contain a match, the "smart" data integration feature goes further by constructing a synthetic data product that represents the complete results of the query both in terms of the attributes and the values that are returned. Each original data package match may have wildly different syntactic structures (e.g. column number, naming, order) but could share a subset of attributes that are asserted as semantically compatible. These compatible, in-common attributes then become the facets and metrics in the synthetic dataset with their values filtered accordingly. While preliminary results have been quite promising - albeit academic - continued development of algorithms for determining compatibility of annotated measurements (e.g. gram and ounce are both mass units) and for mutating measurement values using ontologically-defined conversions (e.g. 1000 milligrams in a gram) will enable more thorough data integration.

>>>>

\subsection{Data query}
For even more precise recall, the OBOE model can be partially \emph{materialized} during the query stage after which a data range filter can be applied. Different techniques are available for merging the Annotation with the data that it describes, but for performance reasons a hybrid approach has been adopted in which premliminary search results from a concept-based query are collated and only that subset is materialized. The Data Manager Library is conveniently used to load raw data objects because our corpus is described using EML and the Annotation syntax - both of which inform the correct use of the Data Manager query features. For any data attributes that match the concept query criteria, we then verify that those attributes contain data values within the range specifed by the initial semantic data query (\figref{fig:metacat-query}).

\subsection{Data integration}
The materialization routine used for semantic data queries laid the groundwork for our ongoing data integration pursuits.
In addition to inspecting the data for values within a range and returning the packages that contain a match, the "smart" data integration feature goes further by constructing a synthetic data product that represents the complete results of the query both in terms of the attributes and the values that are returned. Each original data package match may have wildly different syntactic structures (e.g. column number, naming, order) but could share a subset of attributes that are asserted as semantically compatible. These compatible, in-common attributes then become the facets and metrics in the synthetic dataset with their values filtered accordingly (\figref{fig:integration}). While preliminary results have been quite promising - albeit academic - continued development of algorithms for determining compatibility of annotated measurements (e.g. gram and ounce are both mass units) and for mutating measurement values using ontologically-defined conversions (e.g. 1000 milligrams in a gram) will enable more thorough data integration.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{images/metacat-query.png}
\caption{Semantic data query web interface. Data packages containing observations of Kelp Wet Mass less than or equal to 5 [grams] are returned. Additional search options and compund query criteria can be specified within the other tabs. Matches can be saved in the data cart for later exploration.}
\label{fig:metacat-query}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{images/integration.png}
\caption{Data integration query across multiple data packages (A, B).  Annotations (C, D) are used to determine semantically equivalent data attributes contained in the data objects (E, F). Attributes 'plot' and 'site' are considered equivalent measurements of the characteristic 'Location'; attributes 'weight' and 'wt' both map to the same characteristic 'Mass'. The Semantic Mediation API utilizes the Data Manager Library to load and query the source data informed by semantic similarities between the structurally disparate data attributes.}
\label{fig:integration}
\end{figure}

### 2011-06-02 docs/pubs/EIM-2011/framework.tex
figures for: integration, metacat query, morpho annotation

\subsection{Morpho editor plugin}
The semantic editor plugin for Morpho provides a front-end to the SMS API and allows metadata editors to build up semantic Annotations for existing data package descriptions. The editor focuses on a simplictic fill-in-the-blank style interface with reusable, searchable, hierachical concept selection widgets. The optional plugin seamlessly integrates with a standard Morpho installation augmenting the existing features to provide semantic query capabilities for locating data packages, marking up data packages with semantic Annotations, and saving Annotations locally or to a shared repository where they can be discovered and explored by a broader audience.

>>>>

\subsection{Morpho editor plugin}
The semantic editor plugin for Morpho provides a front-end to the SMS API and allows metadata editors to build up semantic Annotations for existing data package descriptions. The editor focuses on a simplictic fill-in-the-blank style interface with reusable, searchable, hierachical concept selection widgets (\figref{fig:morpho-annotation}). The optional plugin seamlessly integrates with a standard Morpho installation augmenting the existing features to provide semantic query capabilities for locating data packages, marking up data packages with semantic Annotations, and saving Annotations locally or to a shared repository where they can be discovered and explored by a broader audience.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{images/morpho-annotation.png}
\caption{Morpho metadata editor with Semantic plugin. Fill-in-the-blank style editing interface to encourage natural language descriptions. A searchable hierarchical browser is used to select concepts from an ontology.}
\label{fig:morpho-annotation}
\end{figure}
